Dominance Regions for Affine Cluster Algebras

### 2022-06-24 affine_dominance.tex
Section 5 edits

  For $\lambda\in\ZZ^n$ and $r\in\ZZ_{>0}$ with $r=ns+\ell$ for $1\le\ell\le n$, the tropical transformation $\phi_{t_r}(\lambda)$ is given by
  \[
    E^{t_{r-1}}_{\ell,\varepsilon_r} \cdots E^{t_{r-\ell}}_{1,\varepsilon_{r-\ell+1}} E^{t_{ns-1}}_{n,\varepsilon_{ns}} \cdots E^{t_1}_{2,\varepsilon_2} E^{t_0}_{1,\varepsilon_1} \lambda
  \]
  for a sequence of signs $\varepsilon_1,\ldots,\varepsilon_r$ depending on $\lambda$.
  Similarly, the tropical transformation $\phi_{t_{-r}}(\lambda)$ is given by
  \[
    E^{t_{-r+1}}_{n-\ell+1,\varepsilon_{-r}} \cdots E^{t_{-r+\ell}}_{n,\varepsilon_{-r+\ell-1}} E^{t_{-ns+1}}_{1,\varepsilon_{-ns}} \cdots E^{t_{-1}}_{n-1,\varepsilon_{-2}} E^{t_0}_{n,\varepsilon_{-1}} \lambda
  \]
  for a sequence of signs $\varepsilon_{-1},\ldots,\varepsilon_{-r}$ depending on $\lambda$.
  \begin{lemma}
    \label{lem:imaginary mutations}
    If $\lambda\in\ZZ^n$ is imaginary, then every sign $\varepsilon_j$ with $j>0$ above is equal to $-$ and every sign $\varepsilon_j$ with $j<0$ above is equal to $+$.
  \end{lemma}
  \begin{proof}
    With $1,\ldots,n$ being sink adapted, these sign choices give the most complicated transformation at each step.
    That should mean that $\lambda$ is sufficiently far from the initial chamber and remains there under all mutations.
  \end{proof}

  For a $\bfg$-vector $\lambda\in\ZZ^n$, write $\tilde\lambda\in\ZZ^{2n}$ for its extension by zeros.
  \begin{lemma}
    Let $\lambda\in\ZZ^n$ be an imaginary $\bfg$-vector.
    For $r\in\ZZ$, write $\phi_{t_r}(\tilde\lambda)=(\lambda'_1,\ldots,\lambda'_{2n})$.
    \begin{itemize}
      \item For any $r\ge0$, we have $\lambda'_{n+s}=0$ for $1\le s\le n$.
      \item For $-n\le r\le -1$, we have $\lambda'_{n+s}=0$ for $1\le s\le n+r$ and \[\lambda'_{n+s}=\lambda_s+\sum\limits_{s<i_1<\cdots<i_m\le n} b_{si_1} \prod\limits_{p=2}^m b_{i_{p-1}i_p}\lambda_{i_m}\]
      for $n+r+1\le s\le n$.
      \item For any $r<-n$, we have 
        \[\lambda'_{n+s}=\lambda_s+\sum\limits_{s<i_1<\cdots<i_m\le n} b_{si_1} \prod\limits_{p=2}^m b_{i_{p-1}i_p}\lambda_{i_m}\]
        for $1\le s\le n$.
    \end{itemize}
  \end{lemma}
  \begin{proof}
    We work by induction on $r\in\ZZ_{\ge0}$.
    For $r=0$, the claim is immediate from the definition of $\tilde\lambda$.
    Following Lemma~\ref{lem:imaginary mutations}, for $r>0$ we have $\varepsilon_r=-$.
    Writing $r=ns+\ell$, from Lemma~\ref{lem:tropical signs} and Lemma~\ref{lem:preprojective exchange matrices} we see that the $\ell$-th column of $\tilde B^{t_{r-1}}$ has $\bfc$-vector part given by $\bfc^{(r-1)}_\ell$ which is a positive vector.
    It follows that the matrix $E^{t_{r-1}}_{\ell,\varepsilon_r}$ acts as the identity in the last $n$ components and the claim for $r$ follows from the claim for $r-1$.

>>>>

  For $\lambda\in\ZZ^n$ and $r\in\ZZ_{>0}$ with $r=ns+\ell$ for $1\le\ell\le n$, the tropical transformation $\phi_{t_r}(\lambda)$ is given by
  \[
    E^{t_{r-1}}_{\ell,\varepsilon_r} \cdots E^{t_{r-\ell}}_{1,\varepsilon_{r-\ell+1}} E^{t_{ns-1}}_{n,\varepsilon_{ns}} \cdots E^{t_1}_{2,\varepsilon_2} E^{t_0}_{1,\varepsilon_1} \lambda
  \]
  for a sequence of signs $\varepsilon_1,\ldots,\varepsilon_r$ depending on $\lambda$.
  Similarly, the tropical transformation $\phi_{t_{-r}}(\lambda)$ is given by
  \[
    E^{t_{-r+1}}_{n-\ell+1,\varepsilon_{-r}} \cdots E^{t_{-r+\ell}}_{n,\varepsilon_{-r+\ell-1}} E^{t_{-ns+1}}_{1,\varepsilon_{-ns}} \cdots E^{t_{-1}}_{n-1,\varepsilon_{-2}} E^{t_0}_{n,\varepsilon_{-1}} \lambda
  \]
  for a sequence of signs $\varepsilon_{-1},\ldots,\varepsilon_{-r}$ depending on $\lambda$.
  \begin{lemma}
    \label{lem:imaginary mutations}
    If $\lambda\in\ZZ^n$ is imaginary, then every sign $\varepsilon_r$ with $r>0$ above is equal to $-$ and every sign $\varepsilon_r$ with $r<0$ above is equal to $+$.
  \end{lemma}
  \begin{proof}
    With $1,\ldots,n$ being sink adapted, these sign choices give the most complicated transformation at each step.
    That should mean that $\lambda$ is sufficiently far from the initial chamber and remains there under all mutations.
  \end{proof}

  Using this we explicitly compute the action of $\phi_{t_r}$ on imaginary $\bfg$-vectors. 
  For a $\bfg$-vector $\lambda\in\ZZ^n$, write $\tilde\lambda\in\ZZ^{2n}$ for its extension by zeros.
  \begin{proposition}
    Suppose $\lambda\in\ZZ^n$ is imaginary.
    For $r\in\ZZ$, write $\phi_{t_r}(\tilde\lambda)=(\lambda'_1,\ldots,\lambda'_{2n})$.
    Then the following hold:
    \begin{itemize}
      \item For $r>0$ with $r=ns+\ell$, we have (this is for $s=0$ so far)
        \begin{equation}
          \label{eq:positive g-vector mutations}
          \lambda'_k=
            \begin{cases} 
              \lambda_k+\sum\limits_{1\le i_1<\cdots<i_m\le \ell} (-1)^m b_{ki_m} \prod\limits_{p=2}^m b_{i_p i_{p-1}}\lambda_{i_1} & \text{if $\ell+1\le k\le n$;}\\
              -\lambda_k-\sum\limits_{1\le i_1<\cdots<i_m\le \ell} (-1)^m b_{ki_m} \prod\limits_{p=2}^m b_{i_pi_{p-1}}\lambda_{i_1} & \text{if $1\le k\le \ell$;}\\
            \end{cases}
        \end{equation}
        and $\lambda'_{n+k}=0$ for $1\le k\le n$.
      \item For $-n\le r\le -1$, we have $\lambda'_{n+k}=0$ for $1\le k\le n+r$ and \[\lambda'_{n+k}=\lambda_k+\sum\limits_{k<i_1<\cdots<i_m\le n} b_{ki_1} \prod\limits_{p=2}^m b_{i_{p-1}i_p}\lambda_{i_m}\]
      for $n+r+1\le k\le n$.
      \item For any $r<-n$, we have 
        \[\lambda'_{n+k}=\lambda_k+\sum\limits_{k<i_1<\cdots<i_m\le n} b_{ki_1} \prod\limits_{p=2}^m b_{i_{p-1}i_p}\lambda_{i_m}\]
        for $1\le k\le n$.
    \end{itemize}

  \end{proposition}
  \begin{lemma}
    Let $\lambda\in\ZZ^n$ be an imaginary $\bfg$-vector.
    For $r\in\ZZ$, write $\phi_{t_r}(\tilde\lambda)=(\lambda'_1,\ldots,\lambda'_{2n})$.
    \begin{itemize}
      \item For any $r\ge0$, we have $\lambda'_{n+s}=0$ for $1\le s\le n$.
      \item For $-n\le r\le -1$, we have $\lambda'_{n+s}=0$ for $1\le s\le n+r$ and \[\lambda'_{n+s}=\lambda_s+\sum\limits_{s<i_1<\cdots<i_m\le n} b_{si_1} \prod\limits_{p=2}^m b_{i_{p-1}i_p}\lambda_{i_m}\]
      for $n+r+1\le s\le n$.
      \item For any $r<-n$, we have 
        \[\lambda'_{n+s}=\lambda_s+\sum\limits_{s<i_1<\cdots<i_m\le n} b_{si_1} \prod\limits_{p=2}^m b_{i_{p-1}i_p}\lambda_{i_m}\]
        for $1\le s\le n$.
    \end{itemize}
  \end{lemma}
  \begin{proof}
    We work by induction on $r\in\ZZ_{\ge0}$.
    For $r=0$, the claim is immediate from the definition of $\tilde\lambda$.
    Following Lemma~\ref{lem:imaginary mutations}, for $r>0$ we have $\varepsilon_r=-$.
    Writing $r=ns+\ell$, from Lemma~\ref{lem:tropical signs} and Lemma~\ref{lem:preprojective exchange matrices} we see that the $\ell$-th column of $\tilde B^{t_{r-1}}$ has $\bfc$-vector part given by $\bfc^{(r-1)}_\ell$ which is a positive vector.
    It follows that the matrix $E^{t_{r-1}}_{\ell,\varepsilon_r}$ acts as the identity in the last $n$ components and the claim for $r$ follows from the claim for $r-1$.
